Heuristics and Really Hard Instances for Subgraph Isomorphism Problems

### 2016-01-14 paper/paper.tex
More writing about other solvers

?? We're looking at the odd stuff in the middle, which isn't near a phase transition, but is hard.

Is this just a weakness of current subgraph isomorphism solvers? In \cref{figure:alt} we repeat the
induced experiments on smaller pattern and target graphs, using different solving techniques.
Although these techniques are not competitive in absolute terms, we wish to see if the same pattern
of behaviour occurs.

Direct pseudo-boolean encoding: variables for each assignment, cardinality constraints, edge and
non-edge constraints. We use the Clasp solver. We see that hard instances remain hard, including
inside the central region, and that the easy satisfiable instances remain easy. ?? Nag James to
write this bit, and possibly kidnap his baby if he isn't cooperative.

Although not presented here, we tried an equivalent SAT encoding with direct encoding of the
cardinality constraints using the Glucose solver, and saw very similar performance.  We also tried
using a mixed integer encoding along similar lines---Gurobi was only able to solve some of the
trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
limit. We do not plot these results.

>>>>

The region in the parameter space where both pattern and target have medium density is far from a
phase transition, but nevertheless contains instances that are hard for current subgraph isomorphism
solvers. We would like to know whether this is due to a weakness in current solvers, or whether
instances in this region are inherently difficult to solve. In \cref{figure:alt} we repeat the
induced experiments on smaller pattern and target graphs, using different solving techniques.
Although these techniques are not competitive in absolute terms, we wish to see if the same pattern
of behaviour occurs.

We create a pseudo-boolean (PB) encoding as follows. For each pattern vertex $v$ and each target
vertex $w$, create a binary variable which takes the value 1 if and only if that $v$ is mapped to
$w$.  Constraints are added to ensure that each pattern vertex maps to exactly one target vertex,
that each target vertex is mapped to by at most one pattern vertex, and that adjacent edges are
mapped to adjacent edges. Because we are solving the induced problem, we also add constraints to
ensure that non-adjacent edges are mapped to non-adjacent edges. We used the Clasp
solver~\citep{gekakaosscsc11a} to solve the pseudo-boolean instances. The instances that are hard
for the Glasgow solver remain hard for the PB solver, including instances inside the central region,
and the easy satisfiable instances remain easy.

We implemented an equivalent SAT encoding with direct encoding of the cardinality constraints. Using
the Glucose solver, we saw very similar performance to the PB encoding. We also implemented a
similar integer program encoding; the Gurobi solver was only able to solve some of the trivial
satisfiable instances, and was almost never able to prove unsatisfiability within the time limit. We
do not plot the results for either of these solvers.
